---
title: "02_AGF_adult_Ahyacinthus_survival_statistics_ bayes"
author: "Alex Macadam"
date: "2023-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
set.seed(123)
library(tidyverse) #for data wrangling
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(brms)
library(broom.mixed)
library(tidybayes)
library(bayesplot)
library(rstanarm)
library(cmdstanr)
library(ggeffects)
library(rstan)
library(DHARMa)
library(ggridges)
library(easystats)     #framework for stats, modelling and visualisation
library(patchwork)
source('helperFunctions.R')

theme_Publication <- function(base_size=14, base_family="helvetica") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
}
```

Read in data
```{r}
survival.hya <- read.csv('data/Adult-survival.csv', header=T) |>
  filter(Species == "A. hyacinthus", Timepoint == "T5") |> #remove other species data
  mutate(Survival=as.numeric(Survival), Bleaching_D=as.numeric(Bleaching_D), TissueArea_cm2=as.numeric(TissueArea_cm2)) |> #format columns
  mutate(Tank = factor(Tank),
                         Position = factor(Position),
                         Stick = factor(Stick),
                         Genotype = factor(Genotype),
                         Temperature = factor(Temperature, levels=c('Ambient', 'Hot')),
                         Reef = factor(Reef, levels=c('Davies','Palms','Arlington','Martin','Wood')),
                         Timepoint = factor(Timepoint),
                         Species = factor(Species)
                        ) # change columns to factors
```

# Fit the model
form
```{r}
survival.form <- bf(Survival | trials(1) ~ Population * Temperature + (1|Tank),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(1.5, 5), class = 'Intercept') +
    prior(normal(1, 10), class = 'b') +
    prior(student_t(5, 0, 2.5), class = 'sd') 
```

run model
```{r}
survival.brm <- brm(survival.form,
                  data = survival.hya,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "cmdstanr"
                  )
```

conditional effects
```{r}
survival.brm %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

rerun model
```{r}
survival.brm2 <- update(survival.brm,
                      sample_prior = 'yes',
                      cores = 3,
                      refresh = 0)
```

updated conditional effects
```{r}
ce_survival_hya<- survival.brm2 %>%
    conditional_effects() %>%
    plot(points = TRUE)
ce_survival_hya
ce_survival_hya<- ce_survival_hya$`Population:Temperature`$data |>
  mutate(Species = "A. hyacinthus")
```

plot conditional effects
```{r}
pd <- position_dodge(width=0.2)
pal <- wes_palette("Zissou1", 5, type = "discrete")
scalesize <- function(x) sprintf("%.1f", x)

#plot survival
survival_hya_plot1<- ce_survival_hya |>
  ggplot(aes(x=Population, y=estimate__, shape=Temperature, color = Temperature)) +
  geom_pointrange(aes(ymin=lower__, ymax=upper__, color=Temperature), position=position_dodge(width=0.2))+
  scale_y_continuous(limits=c(0,1), breaks=seq(0,1,0.2), labels = scalesize)+
  scale_color_manual(values=pal[c(1,5)]) +
  labs(x='Reef Cross', y="Survival (%)")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
  #      legend.margin=margin(0,0,0,0),
  #      legend.box.margin=margin(0,0,0,0),
  #      plot.margin = margin(rep(0.5,4)),
  #      legend.position="right",
  #      legend.direction = "vertical",
        axis.text.x = element_text(angle = 90)
  #      axis.text.y = element_text(angle = 90)
        )
survival_hya_plot1
```

View prior and posterior
```{r}
survival.brm2 %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm2$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm2$fit |> stan_ac()
```

Rhat
```{r}
survival.brm2$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm2$fit |> stan_ess()
```

Density overlay
```{r}
survival.brm2 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm2, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))
```

Summarise model
back transform from logit scale: by exponentiating on the odds scale. comparing to 1
```{r}
survival.brm2 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 1),
    Pg = ~mean(.x > 1),
    rhat,
    ess_bulk,
    ess_tail
    ) |>
  knitr::kable()
```

Probability scale
```{r}
survival.brm2 |>
  emmeans(~Population * Temperature, type = "response")
```

Planned contrast:
Arlington
Davies
Martin
Palms
Wood

```{r}
cmat=cbind(
  "Central vs North" = c(0, 0, 0, 0, 0, -1/3, -1/3, 1/2, -1/3, 1/2),
  "Arlington" = c(-1, 0, 0, 0, 0, 1, 0, 0, 0, 0),
  "Davies" = c(0, -1, 0, 0, 0, 0, 1, 0, 0, 0),
  "Martin" = c(0, 0, -1, 0, 0, 0, 0, 1, 0, 0),
  "Palms" = c(0, 0, 0, -1, 0, 0, 0, 0, 1, 0),
  "Wood" = c(0, 0, 0, 0, -1, 0, 0, 0, 0, 1)
)

survival.em<- survival.brm2 |>
  emmeans(~Population * Temperature, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```

#Plot
get variables
```{r}
survival.brm2 %>% get_variables()
```

```{r}
survival.brm2 %>%
  gather_draws(`b_Intercept.*|b_Population.*`, regex = TRUE) %>%
  ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = .variable,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE)
```

####
Hot only
```{r}
survival_hot <- survival.hya |>
  filter(Temperature == "Hot")
```

# Fit the model
form
```{r}
survival.form <- bf(Survival | trials(1) ~ Population + (1|Tank),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(0, 1.7), class = 'Intercept') +
    prior(normal(0, 1.7), class = 'b') +
    prior(student_t(3, 0, 1.7), class = 'sd') 
```

run model- n
```{r}
survival.brm3 <- brm(survival.form, 
                  data = survival_hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "cmdstanr"
                  )
```

conditional effects
```{r}
survival.brm3 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

View priors and posteriors
```{r}
survival.brm3 %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm3$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm3$fit |> stan_ac()
```

Rhat
```{r}
survival.brm3$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm3$fit |> stan_ess()
```

Density overlay
```{r}
survival.brm3 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm3, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))
```

Summarise model
back transform from logit scale: by exponentiating on the odds scale. comparing to 1
```{r}
survival.brm3 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 1),
    Pg = ~mean(.x > 1),
    rhat,
    ess_bulk,
    ess_tail
    ) |>
  knitr::kable()
```

Probability scale
```{r}
survival.brm3 |>
  emmeans(~Population, type = "response")
```

Planned contrast:
Arlington
Davies
Martin
Palms
Wood
```{r}
cmat=cbind(
  "Central vs North" = c(-1/3, -1/3, 1/2, -1/3, 1/2)
)

survival.em<- survival.brm3 |>
  emmeans(~Population, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```

#Plot
get variables
```{r}
survival.brm3 %>% get_variables()
```

```{r}
surv.plot<- survival.brm3 %>%
  emmeans(~Population) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = Population,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```