---
title: "09_AGF_larvae_Gretiformis_survival_statistics_ bayes"
author: "Alex Macadam"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
set.seed(123)
library(tidyverse) #for data wrangling
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(brms)
library(broom.mixed)
library(tidybayes)
library(bayesplot)
library(rstanarm)
library(cmdstanr)
library(ggeffects)
library(rstan)
library(DHARMa)
library(ggridges)
library(easystats)
library(patchwork)
source('helperFunctions.R')
```

```{r}
#load data
survival.hot <- read.csv("data/Larvae-survival-2020.csv", header=T, fileEncoding="UTF-8-BOM") |>
  mutate(Family = factor(Family),
         Temp = factor(Temp),
         Treatment = factor(Treatment),
         Cone = factor(Cone),
         CrossType = factor(CrossType),
         BoatID = factor(BoatID),
         Cone = factor(Cone),
         Culture = factor(Culture),
         popcross = factor(popcross),
         BoatID = factor(BoatID)
         ) |>
  filter(Timepoint == "2", Year==2020, Species == "G.ret", Treatment == "Hot") |> 
  drop_na(No.LarvaeIn)  # drop nas from the start
```

# Fit the model
form
```{r}
survival.form <- bf(Survival | trials(1) ~ popcross + (1|BoatID),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(0, 1.5), class = 'Intercept') +
    prior(normal(0, 1.5), class = 'b') +
    prior(student_t(3, 0, 1.5), class = 'sd') 
```

run model
```{r}
survival.brm <- brm(survival.form, 
                  data = survival.hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "cmdstanr"
                  )
```
View prior and posterior
```{r}
survival.brm %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm$fit |> stan_ac()
```

Rhat
```{r}
survival.brm$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm$fit |> stan_ess()
```

Denisty overlay
```{r}
survival.brm |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))

survival.resids<- make_brms_dharma_res(survival.brm, integerResponse = FALSE)
testUniformity(survival.resids)
plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))
plotResiduals(survival.resids, quantreg = FALSE)
testDispersion(survival.resids)
```

Summarise model
conditional effects
```{r}
survival.brm %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

back transform from logit scale: by exponentiating on the odds scale. comparing to 1
```{r}
survival.brm |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 1),
    Pg = ~mean(.x > 1),
    rhat,
    ess_bulk,
    ess_tail
    ) |>
  knitr::kable()
```

Probability scale
```{r}
survival.brm |>
  emmeans(~popcross, type = "response")
```

Planned contrast:
                            Jewmum_vs_Swimum |   JewPar_vs_Swimum  | Jewmum_vs_Jewdad
JewellBulkxParkeBulk               -1/2            -1/4                     -1/2                   
JewellBulkxSwitzerBulk             -1/2            -1/4                     -1/2                 
ParkeBulkxJewellBulk                0              -1/4                      1/2                   
ParkeBulkxSwitzerBulk               0              -1/4                      0                  
SwitzerBulkxJewellBulk              1/2             1/2                      1/2                     
SwitzerBulkxParkeBulk               1/2             1/2                      0                     

```{r}
cmat=cbind(
  "Jewmum_vs_Swimum" = c(-0.5, -0.5, 0, 0, 0.5, 0.5),
  "JewPar_vs_Swimum" = c(-0.25,-0.25,-0.25,-0.25,0.5,0.5),
  "Jewmum_vs_Jewdad" = c(-0.5,-0.5,0.5,0,0.5,0)
)
```

```{r}
survival.em<- survival.brm |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```

#Plot
```{r}
surv.plot<- survival.brm %>%
  emmeans(~popcross) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = popcross,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```

######
By region cross
```{r}
survival.hot <- survival.hot %>%
  mutate(Mom_cross = case_when(
    popcross %in% c("ArlingtonxArlington", "ArlingtonxDavies", "DaviesxArlington", "DaviesxDavies") ~ "central",
    popcross %in% c("DaviesxMartin", "MartinxArlington", "MartinxDavies", "MartinxPalms", 
                    "PalmsxMartin", "WoodxArlington") ~ "hybrid",
    popcross %in% c("MartinxMartin", "MartinxWood", "WoodxMartin") ~ "north",
    TRUE ~ NA_character_  # Assign NA if no match
  ))
```

form
```{r}
survival.form <- bf(Survival | trials(1) ~ Mom + (1|BoatID),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(0, 1.5), class = 'Intercept') +
    prior(normal(0, 1.5), class = 'b') +
    prior(student_t(3, 0, 1.5), class = 'sd') 
```

run model
```{r}
survival.brm4 <- brm(survival.form, 
                  data = survival.hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "cmdstanr"
                  )
```

View prior and posterior
```{r}
survival.brm4 %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm4$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm4$fit |> stan_ac()
```

Rhat
```{r}
survival.brm4$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm4$fit |> stan_ess()
```

Denisty overlay
```{r}
survival.brm4 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))

survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
testUniformity(survival.resids)
plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))
plotResiduals(survival.resids, quantreg = FALSE)
testDispersion(survival.resids)
```

Summarise model
conditional effects
```{r}
survival.brm4 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

```{r}
surv.plot<- survival.brm4 %>%
  emmeans(~Mom) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = Mom,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```
