---
title: "10_AGF_juvenile_Akenti_survival_statistics_ bayes"
author: "Alex Macadam"
date: "2023-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
set.seed(123)
library(tidyverse) #for data wrangling
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(brms)
library(broom.mixed)
library(tidybayes)
library(bayesplot)
library(rstanarm)
library(cmdstanr)
library(ggeffects)
library(rstan)
library(DHARMa)
library(ggridges)
library(easystats)
library(patchwork)
source('helperFunctions.R')
```

```{r}
#load data
survival <- read.csv("data/Juvenile-survival-2020.csv", header=T, fileEncoding="UTF-8-BOM") |>
  mutate(Family = factor(Family),
         Species = factor(Species),
         Temp = factor(Temp),
         Tank = factor(Tank),
         Timepoint = factor(Timepoint),
         Cassette = factor(Cassette),
         Plate.Slide = factor(Plate.Slide),
         Treatment = factor(Treatment),
         Cone = factor(Cone),
         Cross = factor(Cross),
         Crosstype = factor(Crosstype),
         RegionCross = factor(RegionCross),
         popcross = factor(popcross)
         ) |>
  filter(Timepoint == "6", Crosstype == "Individual Cross")

survival.hot <- survival |> filter(Temp == "HOT")
```

# Fit the model
form
```{r}
survival.form <- bf(Survival | trials(1) ~ popcross + (1|Tank),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(1, 1.7), class = 'Intercept') +
    prior(normal(0, 1.7), class = 'b') +
    prior(student_t(3, 0, 1.7), class = 'sd') 
```

run model
```{r}
survival.brm <- brm(survival.form, 
                  data = survival.hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "cmdstanr"
                  )
```

View prior and posterior
```{r}
survival.brm %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm$fit |> stan_ac()
```

Rhat
```{r}
survival.brm$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm$fit |> stan_ess()
```

Denisty overlay
```{r}
survival.brm |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))

survival.resids<- make_brms_dharma_res(survival.brm, integerResponse = FALSE)
testUniformity(survival.resids)
plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))
plotResiduals(survival.resids, quantreg = FALSE)
testDispersion(survival.resids)
```

Summarise model
conditional effects
```{r}
survival.brm %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

back transform from logit scale: by exponentiating on the odds scale. comparing to 1
```{r}
survival.brm |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 1),
    Pg = ~mean(.x > 1),
    rhat,
    ess_bulk,
    ess_tail
    ) |>
  knitr::kable()
```

Probability scale
```{r}
survival.brm |>
  emmeans(~popcross, type = "response")
```

Planned contrast:
 DaviesxDavies
 JewellxDavies
 JewellxJewell
 JewellxPalms
 PalmsxDavies
 PalmsxPalms
 ParkexDavies
 ParkexParke
 SwitzerxDavies
 SwitzerxSwitzer
```{r}
cmat=cbind(
  "PureSouth vs Hybrid" = c(-1/3, 1/4, 0, 1/4, -1/3, -1/3, 1/4, 0, 1/4, 0),
  "PureNorth vs Hybrid" = c(0, 1/4, -1/4, 1/4, -1/4, 0, 1/4, -1/4, 1/4, -1/4),
  "Central v North" = c(-1/3, 0, 1/3, 0, -1/3, -1/3, 0, 1/3, 0, 1/3),
  "Dav vs JewDav" = c(-1, 1, 0, 0, 0, 0, 0, 0, 0, 0),
  "Pal vs JewPal" = c(0, 0, 0, 1, 0, -1, 0, 0, 0, 0),
  "Dav vs ParDav" = c(-1, 0, 0, 0, 0, 0, 1, 0, 0, 0),
  "Dav vs SwiDav" = c(-1, 0, 0, 0, 0, 0, 0, 0, 1, 0)
)
```

```{r}
survival.em<- survival.brm4 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```

#Plot
```{r}
surv.plot<- survival.brm %>%
  emmeans(~popcross) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = popcross,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```

######
By Symbionts 
form
```{r}
survival.form <- bf(Survival | trials(1) ~ Treatment * Temp + (1|Tank),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(2, 1.7), class = 'Intercept') +
    prior(normal(0, 1.7), class = 'b') +
    prior(student_t(3, 0, 1.7), class = 'sd') 
```

run model
```{r}
survival.brm3 <- brm(survival.form, 
                  data = survival,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 100,
                  backend = "cmdstanr"
                  )
```

View prior and posterior
```{r}
survival.brm3 %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm3$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm3$fit |> stan_ac()
```

Rhat
```{r}
survival.brm3$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm3$fit |> stan_ess()
```

Denisty overlay
```{r}
survival.brm3 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm3, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))

survival.resids<- make_brms_dharma_res(survival.brm3, integerResponse = FALSE)
testUniformity(survival.resids)
plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))
plotResiduals(survival.resids, quantreg = FALSE)
testDispersion(survival.resids)
```

Summarise model
conditional effects
```{r}
survival.brm3 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

Probability scale
```{r}
survival.brm3 |>
  emmeans(~Treatment * Temp, type = "response")
```

Planned contrast:
D1 amb
Sed amb
SS amb
D1 hot
Sed hot
SS hot

```{r}
cmat=cbind(
  "D1" = c(-1, 0, 0, 1, 0, 0),
  "SS" = c(0, -1, 0, 0, 1, 0),
  "Sed" = c(0, 0, -1, 0, 0, 1)
)
```

```{r}
survival.em<- survival.brm3 |>
  emmeans(~Treatment * Temp, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```

```{r}
surv.plot<- survival.brm3 %>%
  emmeans(~Treatment * Temp) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = Treatment,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```

####
By Region cross
form
```{r}
survival.form <- bf(Survival | trials(1) ~ RegionCross + (1|Tank),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(0, 1.5), class = 'Intercept') +
    prior(normal(0, 1.5), class = 'b') +
    prior(student_t(3, 0, 1.5), class = 'sd') 
```

run model- n
```{r}
survival.brm4 <- brm(survival.form,
                  data = survival_hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "cmdstanr"
                  )
```

View prior and posterior
```{r}
survival.brm4 %>% SUYR_prior_and_posterior()
```

Trace plots
```{r}
survival.brm4$fit |> stan_trace()
```

Autocorrelation
```{r}
survival.brm4$fit |> stan_ac()
```

Rhat
```{r}
survival.brm4$fit |> stan_rhat()
```

Effective sample size
```{r}
survival.brm4$fit |> stan_ess()
```

Denisty overlay
```{r}
survival.brm4 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))

survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
testUniformity(survival.resids)
plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))
plotResiduals(survival.resids, quantreg = FALSE)
testDispersion(survival.resids)
```

Summarise model
conditional effects
```{r}
survival.brm4 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

```{r}
surv.plot<- survival.brm4 %>%
  emmeans(~RegionCross) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = RegionCross,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```