---
title: "Physiology- A. hyacinthus larvae survival statistics"
author: "Alex Macadam"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Survival statistics
```{r}
library(tidyverse) #for data wrangling
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(brms)
library(broom.mixed)
library(tidybayes)
library(bayesplot)
#library(standist)   #for visualizing distributions
library(rstanarm)
library(ggeffects)
library(rstan)
library(DHARMa)
library(ggridges)
library(easystats)     #framework for stats, modelling and visualisation
library(patchwork)
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(broom.mixed)
library(ggfortify) #for model diagnostics
#library(sjPlot)    #for outputs
library(knitr)     #for kable
source('/Users/alexmacadam/Dropbox/PhD/Chapter5_Outplanting/Chapter 5 Scripts/Chapter-5--Outplanting-AGF-corals/helperFunctions.R')
```

```{r}
#load data
survival <- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter2_Physiology/Chapter 2 data/Bayes stats data/survival/larvae_surv2021.csv", header=T, fileEncoding="UTF-8-BOM")
```

```{r}
#change catagorical columns to factors and remove NAs
survival <- survival |>
  mutate(Family = factor(Family),
         Temp = factor(Temp),
         Treatment = factor(Treatment),
         Cone = factor(Cone),
         CrossType = factor(CrossType),
         BoatID = factor(BoatID),
         Cone = factor(Cone),
         Culture = factor(Culture),
         popcross = factor(popcross),
         BoatID = factor(BoatID)
         ) |>
  filter(Timepoint == "2") |> 
  drop_na(No.LarvaeIn)  # drop nas from the start
```

```{r}
#survival |>
#ggplot(aes(y=Survival, x=Cross2)) +
#    geom_point(position=position_jitter(width=0.2, height=0)) +
#    facet_wrap(~RegionCross)
```

# Fit the model
## Random intercept
form
```{r}
#survival.form <- bf(Survival | trials(1) ~ popcross * Temp + (1|BoatID),
#                  family=binomial(link='logit'))
```

```{r}
options(width=150)
survival.form %>% get_prior(data = survival)
options(width=80)
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
#priors <-
#    prior(normal(0, 1.5), class = 'Intercept') +
#    prior(normal(0, 1.5), class = 'b') +
#    prior(student_t(3, 0, 1.5), class = 'sd') 
```

run model- n
```{r}
#survival.brm <- brm(survival.form, 
#                  data = survival,
#                  prior = priors,
#                  sample_prior = 'only',
#                  iter = 5000,
#                  warmup = 2500,
#                  chains = 3, cores = 3,
#                  thin = 5,
#                  control = list(adapt_delta = 0.99, max_treedepth = 20),
#                  refresh = 0,
#                  backend = "rstan"
#                  )
```

conditional effects
```{r}
#survival.brm %>%
#    conditional_effects() %>%
#    plot(points = TRUE)
```

rerun model
```{r}
#survival.brm2 <- update(survival.brm,
#                      sample_prior = 'yes',
#                      cores = 3,
#                      refresh = 0)
```

updated conditional effects
```{r}
#survival.brm2 %>%
#    conditional_effects() %>%
#    plot(points = TRUE)
```

```{r}
#survival.brm2 %>% SUYR_prior_and_posterior()
```

```{r}
#survival.brm2$fit |> stan_trace()
```

```{r}
#survival.brm2$fit |> stan_ac()
```

```{r}
#survival.brm2$fit |> stan_rhat()
```

```{r}
#survival.brm2$fit |> stan_ess()
```

##Random slope model

priors
```{r}
priors <-
    prior(normal(0, 1.7), class = 'Intercept') +
    prior(normal(0, 3), class = 'b') +
    prior(student_t(3,0,1.7), class = 'sd', coef = "Intercept", group = "BoatID") +
    prior(student_t(3, 0, 1.7), class = 'sd') +
    prior(lkj_corr_cholesky(1), class = 'cor')
```

form
```{r}
survival.form <- bf(Survival | trials(1) ~ popcross * Treatment + (popcross|BoatID),
                  family=binomial(link='logit'))
```

```{r}
#survival.brm3 <- brm(survival.form, 
#                  data = survival,
#                  prior = priors,
#                  sample_prior = 'yes',
#                  iter = 5000,
#                  warmup = 2500,
#                  chains = 3, cores = 3,
#                  thin = 5,
#                  control = list(adapt_delta = 0.99, max_treedepth = 20),
#                  refresh = 0,
#                  backend = "rstan"
#                  )
```

```{r}
#saveRDS(survival.brm3, file = "survival.brm3_A.hyacinthus.rds")
survival.brm3 <- readRDS("survival.brm3_A.hyacinthus.rds")
```

```{r}
survival.brm3 %>% SUYR_prior_and_posterior()
```

```{r}
survival.brm3$fit |> stan_trace()
```

```{r}
survival.brm3$fit |> stan_ac()
```

```{r}
survival.brm3$fit |> stan_rhat()
```

```{r}
survival.brm3$fit |> stan_ess()
```

##Compare models
compare loos
```{r}
#(l.1 <- survival.brm2 %>% loo())
#(l.2 <- survival.brm3 %>% loo())
#loo_compare(l.1, l.2)
#loo(survival.brm2, survival.brm3)
```

```{r}
survival.brm3 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm3, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))
```

Summarise model
conditional effects
```{r}
survival.brm3 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

back transform from logit scale: by exponentiating on the odds scale. comparing to 1
```{r}
survival.brm3 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 1),
    Pg = ~mean(.x > 1),
    rhat,
    ess_bulk,
    ess_tail
    ) |>
  knitr::kable()
```

Probability scale
```{r}
survival.brm3 |>
  emmeans(~popcross, type = "response")
```

Planned contrast:

                      Pure vs Hybrid | PureSouth vs Hybrid | PureNorth vs Hybrid | Pure Dav vs hybrid Dav | Mart mom vs Mart dad | Hot vs cool reef
ArlingtonxArlington          -1/3             -1/2                      0              0                         0                   1/2
ArlingtonxDavies              1/10             1/7                      1/8            0                         0                   0
DaviesxArlington              1/10             1/7                      1/8            0                         0                   0
DaviesxDavies                -1/3             -1/2                      0             -1                         0                   1/2
DaviesxMartin                 1/10             1/7                      1/8            1/2                       1/2                 0
MartinxArlington              1/10             1/7                      1/8            0                        -1/2                 0
MartinxDavies                 1/10             1/7                      1/8            1/2                      -1/2                 0
MartinxMartin                -1/3              0                       -1/3            0                         0                  -1/3
MartinxPalms                  1/10             1/7                      1/8            0                        -1/2                 0
MartinxWood                   1/10             0                       -1/3            0                        -1/2                -1/3
PalmsxMartin                  1/10             1/7                      1/8            0                         1/2                 0
WoodxArlington                1/10             0                        1/8            0                         0                   0
WoodxMartin                   1/10             0                       -1/3            0                         0                  -1/3

```{r}
cmat=cbind(
  "Pure vs Hybrid" = c(-1/3, 1/10, 1/10, -1/3, 1/10, 1/10, 1/10, -1/3, 1/10, 1/10, 1/10, 1/10, 1/10),
  "PureSouth vs Hybrid" = c(1/2, -1/7, -1/7, 1/2, -1/7, -1/7, -1/7, 0, -1/7, 0, -1/7, 0, 0),
  "PureNorth vs Hybrid" = c(0, 1/8, 1/8, 0, 1/8, 1/8, 1/8, -1/3, 1/8, -1/3, 1/8, 1/8, -1/3),
  "Pure Dav vs hybrid Dav" = c(0, 0, 0, -1, 1/2, 0, 1/2, 0, 0, 0, 0, 0, 0),
  "Mart mom vs Mart dad" = c(0, 0, 0, 0, 1/2, -1/2, -1/2, 0, -1/2, -1/2, 1/2, 0, 0),
  "Hot vs cool reef" = c(-1/2, 0, 0, -1/2, 0, 0, 0, 1/3, 0, 1/3, 0, 0, 1/3),
  "Dav vs ArlDav" = c(0, 1, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
  "Dav vs MarDav" = c(0, 0, 0, -1, 0, 0, 1, 0, 0, 0, 0, 0, 0),
  "Arl vs MarArl" = c(-1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0),
  "North vs South" = c(-1/4, -1/4, -1/4, -1/4, 0, 0, 0, 1/3, 0, 1/3, 0, 0, 1/3)
)
```

```{r}
survival.em <- survival.brm3 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat))
survival.em
```
odds scale:
```{r}
survival.brm3 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws()
#full contrasts on the link scale
```

```{r}
survival.em<- survival.brm3 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```
Hot vs cool reef: hot are x more likely to survive than cool. strong evidence
Mart mom vs Mart dad: martin mom larvae are x more likely to survive than martin dads. strong evidence
Pure Dav vs hybrid Dav: pure davies larvae are x more likely to survive than hybrid davies. evidence
Pure vs Hybrid: hybrid are not more likely to survive than pure larvae.
PureNorth vs Hybrid: pure norths are x more likely to survive than hybrids. strong evidence
PureSouth vs Hybrid: pure souths are not more likely to survive than hybrids.

#Plot
get variables
```{r}
survival.brm3 %>% get_variables()
```

```{r}
survival.draw <- survival.brm3 %>%
  gather_draws(`b.Intercept.*|b_popcross.*`, regex = TRUE)
survival.draw
```

```{r}
survival.draw %>% median_hdci()
```

```{r}
survival.draw %>%
  mutate(.value = exp(.value)) %>%
  median_hdci()
```

```{r}
survival.brm3 %>%
  gather_draws(`b_Intercept.*|b_popcross.*`, regex = TRUE) %>%
  ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = .variable,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE)
```

```{r}
## Or in colour
survival.brm3 %>%
  gather_draws(`^b_.*`, regex = TRUE) %>%
  filter(.variable != "b_Intercept") %>%
  ggplot() +
  geom_density_ridges_gradient(aes(
    x = (.value),
    y = .variable,
    fill = stat(x)
  ),
  alpha = 0.4, colour = "white",
  quantile_lines = TRUE,
  quantiles = c(0.025, 0.975)
  ) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_continuous() +
  scale_fill_viridis_c(option = "C")
```

##################################################################
Hot only
```{r}
survival_hot<- survival |>
  filter(Treatment == "Hot")
```


priors
```{r}
priors <-
    prior(normal(0, 1.7), class = 'Intercept') +
    prior(normal(0, 3), class = 'b') +
    prior(student_t(3,0,1.7), class = 'sd', coef = "Intercept", group = "BoatID") +
    prior(student_t(3, 0, 1.7), class = 'sd') +
    prior(lkj_corr_cholesky(1), class = 'cor')
```

form
```{r}
survival.form <- bf(Survival | trials(1) ~ popcross + (popcross|BoatID),
                  family=binomial(link='logit'))
```

```{r}
survival.brm4 <- brm(survival.form, 
                  data = survival_hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "rstan"
                  )
```

```{r}
#saveRDS(survival.brm4, file = "survival.brm4_A.hyacinthus.rds")
survival.brm4 <- readRDS("survival.brm4_A.hyacinthus.rds")
```

conditional effects
```{r}
survival.brm4 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

```{r}
survival.brm4 %>% SUYR_prior_and_posterior()
```

```{r}
survival.brm4$fit |> stan_trace()
```

```{r}
survival.brm4$fit |> stan_ac()
```

```{r}
survival.brm4$fit |> stan_rhat()
```

```{r}
survival.brm4$fit |> stan_ess()
```

```{r}
survival.brm4 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))
```

Summarise model
back transform from logit scale: by exponentiating on the odds scale. comparing to 1
```{r}
survival.brm4 |>
  as_draws_df() |>
  mutate(across(everything(), exp)) |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 1),
    Pg = ~mean(.x > 1),
    rhat,
    ess_bulk,
    ess_tail
    ) |>
  knitr::kable()
```

Probability scale
```{r}
survival.brm4 |>
  emmeans(~popcross, type = "response")
```

Planned contrast:

                      Pure vs Hybrid | PureSouth vs Hybrid | PureNorth vs Hybrid | Pure Dav vs hybrid Dav | Mart mom vs Mart dad | Hot vs cool reef
ArlingtonxArlington          -1/3             -1/2                      0              0                         0                   1/2
ArlingtonxDavies              1/10             1/7                      1/8            0                         0                   0
DaviesxArlington              1/10             1/7                      1/8            0                         0                   0
DaviesxDavies                -1/3             -1/2                      0             -1                         0                   1/2
DaviesxMartin                 1/10             1/7                      1/8            1/2                       1/2                 0
MartinxArlington              1/10             1/7                      1/8            0                        -1/2                 0
MartinxDavies                 1/10             1/7                      1/8            1/2                      -1/2                 0
MartinxMartin                -1/3              0                       -1/3            0                         0                  -1/3
MartinxPalms                  1/10             1/7                      1/8            0                        -1/2                 0
MartinxWood                   1/10             0                       -1/3            0                        -1/2                -1/3
PalmsxMartin                  1/10             1/7                      1/8            0                         1/2                 0
WoodxArlington                1/10             0                        1/8            0                         0                   0
WoodxMartin                   1/10             0                       -1/3            0                         0                  -1/3

```{r}
cmat=cbind(
  "Pure vs Hybrid" = c(-1/3, 1/10, 1/10, -1/3, 1/10, 1/10, 1/10, -1/3, 1/10, 1/10, 1/10, 1/10, 1/10),
  "PureSouth vs Hybrid" = c(-1/2, 1/7, 1/7, -1/2, 1/7, 1/7, 1/7, 0, 1/7, 0, 1/7, 0, 0),
  "PureNorth vs Hybrid" = c(0, 1/8, 1/8, 0, 1/8, 1/8, 1/8, -1/3, 1/8, -1/3, 1/8, 1/8, -1/3),
  "Pure Dav vs hybrid Dav" = c(0, 0, 0, -1, 1/2, 0, 1/2, 0, 0, 0, 0, 0, 0),
  "Mart mom vs Mart dad" = c(0, 0, 0, 0, 1/2, -1/2, -1/2, 0, -1/2, -1/2, 1/2, 0, 0),
  "Hot vs cool reef" = c(1/2, 0, 0, 1/2, 0, 0, 0, -1/3, 0, -1/3, 0, 0, -1/3)
)
```

```{r}
survival.em <- survival.brm4 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat))
survival.em
```
odds scale:
```{r}
survival.brm4 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws()
#full contrasts on the link scale
```

```{r}
survival.em<- survival.brm4 |>
  emmeans(~popcross, type = 'response') |>
  contrast(method=list(cmat)) |>
  gather_emmeans_draws() |>
  mutate(across(everything(), exp)) |>
#on odds ratio scale
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)
            )
survival.em
```

#Plot
get variables
```{r}
survival.brm4 %>% get_variables()
```

```{r}
survival.draw <- survival.brm4 %>%
  gather_draws(`b.Intercept.*|b_popcross.*`, regex = TRUE)
survival.draw
```

```{r}
survival.draw %>% median_hdci()
```

```{r}
survival.draw %>%
  mutate(.value = exp(.value)) %>%
  median_hdci()
```

```{r}
survival.brm4 %>%
  gather_draws(`b_Intercept.*|b_popcross.*`, regex = TRUE) %>%
  ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = .variable,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE)
```

```{r}
## Or in colour
survival.brm4 %>%
  gather_draws(`^b_.*`, regex = TRUE) %>%
  filter(.variable != "b_Intercept") %>%
  ggplot() +
  geom_density_ridges_gradient(aes(
    x = (.value),
    y = .variable,
    fill = stat(x)
  ),
  alpha = 0.4, colour = "white",
  quantile_lines = TRUE,
  quantiles = c(0.025, 0.975)
  ) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_continuous() +
  scale_fill_viridis_c(option = "C")
```

######
By region cross
```{r}
survival_hot <- survival_hot %>%
  mutate(Region_cross = case_when(
    popcross %in% c("ArlingtonxArlington", "ArlingtonxDavies", "DaviesxArlington", "DaviesxDavies") ~ "central",
    popcross %in% c("DaviesxMartin", "MartinxArlington", "MartinxDavies", "MartinxPalms", 
                    "PalmsxMartin", "WoodxArlington") ~ "hybrid",
    popcross %in% c("MartinxMartin", "MartinxWood", "WoodxMartin") ~ "north",
    TRUE ~ NA_character_  # Assign NA if no match
  ))
```

# Fit the model
## Random intercept
form
```{r}
survival.form <- bf(Survival | trials(1) ~ Region_cross + (1|BoatID),
                  family=binomial(link='logit'))
```

priors
```{r fitModel2h1, results='markdown', eval=TRUE, mhidden=TRUE, cache=TRUE}
priors <-
    prior(normal(0, 1.5), class = 'Intercept') +
    prior(normal(0, 1.5), class = 'b') +
    prior(student_t(3, 0, 1.5), class = 'sd') 
```

run model- n
```{r}
survival.brm4 <- brm(survival.form, 
                  data = survival_hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 2500,
                  chains = 3, cores = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0,
                  backend = "rstan"
                  )
```

```{r}
survival.brm4 %>% SUYR_prior_and_posterior()
```

```{r}
survival.brm4$fit |> stan_trace()
```

```{r}
survival.brm4$fit |> stan_ac()
```

```{r}
survival.brm4$fit |> stan_rhat()
```

```{r}
survival.brm4$fit |> stan_ess()
```

```{r}
survival.brm4 |> pp_check(type = 'dens_overlay', nsamples = 250)
```

Dharma residuals
```{r}
survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
wrap_elements(~testUniformity(survival.resids)) +
wrap_elements(~plotResiduals(survival.resids, form = factor(rep(1,nrow(survival))))) +
wrap_elements(~plotResiduals(survival.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(survival.resids))

survival.resids<- make_brms_dharma_res(survival.brm4, integerResponse = FALSE)
testUniformity(survival.resids)
plotResiduals(survival.resids, form = factor(rep(1,nrow(survival.hot))))
plotResiduals(survival.resids, quantreg = FALSE)
testDispersion(survival.resids)
```

Summarise model
conditional effects
```{r}
survival.brm4 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

```{r}
surv.plot<- survival.brm4 %>%
  emmeans(~Region_cross) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = Region_cross,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Survival", y = "Treatment")
surv.plot
```