---
title: "Physiology- A. hyacinthus juvenile size"
author: "Alex Macadam"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the necessary libraries
```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(tidyverse) #for data wrangling
library(brms)
library(tidybayes)
library(broom.mixed)#for tidying MCMC outputs
library(patchwork)  #for multiple plots
library(rstanarm)
library(ggeffects)
library(bayesplot)
library(rstan)
library(DHARMa)
library(ggridges)
source('C:/Users/amacadam/Dropbox/PhD/Chapter 5- Outplanting/Chapter 5 Scripts/Chapter-5--Outplanting-AGF-corals/helperFunctions.R')
```

# Read in the data
```{r readData, results='markdown', eval=TRUE}
size <- read.csv("C:/Users/amacadam/Dropbox/PhD/Chapter 2- Physiology/Chapter 2 data/Bayes stats data/physiology/juv_size2021.csv", header=T, fileEncoding="UTF-8-BOM")
```

# Exploratory data analysis
```{r processData, results='markdown', eval=TRUE}
#change catagorical columns to factors and remove NAs
size <- size |>
  mutate(Family = factor(Family),
         Species = factor(Species),
         Temp = factor(Temp),
         Tank = factor(Tank.No.),
         Timepoint = factor(Timepoint),
         Cassette = factor(Cassette),
         Plate.Slide = factor(Slide),
         Treatment = factor(Treatment),
         Cone = factor(Cone),
         Cross = factor(Cross),
         CrossType = factor(CrossType),
         popcross = factor(popcross),
         SINGLE.CLUMP = factor(SINGLE.CLUMP),
         Mom = factor(Mom),
         Dad = factor(Dad)
         ) |>
  filter(CrossType == "IndivCross", SINGLE.CLUMP == "S", Timepoint != "3")
```

Visualise
```{r}
ggplot(size, aes(y = Area, x = Treatment, colour = Temp)) +
  geom_point(position = position_jitter(width = 0.2, height = 0)) +
  facet_wrap(~popcross)
```

```{r eda1, results='markdown', eval=TRUE, hidden=TRUE}
ggplot(size, aes(y=Area, x=popcross)) +
    geom_boxplot() +
    facet_grid(~Treatment)
## Better still
ggplot(size, aes(y=Area, x=Temp, group=Treatment)) +
    geom_point() +
    geom_line() +
    facet_grid(~popcross) 
```

Add Temp x Treatment column
```{r}
size$popcrossxsym <- paste(size$Treatment, size$popcross, sep='_')
size$popcrossxsym <- as.factor(size$popcrossxsym)
```

```{r}
growth<- size %>% filter(Timepoint2 %in% c("T0","T6")) %>% dplyr::select(Timepoint2, Temp,  Cross, Area, IndID, popcross, Tank, Cassette, Plate.Slide, Mom, Dad) %>%
  mutate(popcross = factor(popcross)) %>%
  pivot_wider(names_from = Timepoint2, values_from = Area) %>%
  tidyr::unnest() %>%
  mutate(Growth = ((`T6` - `T0`)/`T0`)*100) |>
  drop_na(Growth)
```

### Using default priors
```{r}
growth.form <- bf(Growth ~ popcross * Temp + (1|Plate.Slide),
                     family = gaussian()
                   )
```

```{r}
options(width=150)
growth.form %>% get_prior(data = growth)
options(width=80)
```

### Defining priors
```{r}
growth |>
group_by(popcross, Temp) |>
summarise(
mean(Growth),
median(Growth),
sd(Growth),
mad(Growth)
)
```

```{r fitModel2h1, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
priors<- prior("normal(-10, 5)", class = "Intercept") +
    prior("normal(0, 15)", class = "b") +
    #prior("normal(0, 0.5)", class = "sd", group = "Tank") +
    prior("normal(0, 20)", class = "sd", group = "Plate.Slide") +
    prior(student_t(3, 0, 24.6), class = 'sigma') +
    prior(cauchy(0,2.5), class = 'sd')
```

```{r}
growth.brm <- brm(growth.form, 
                  data = growth,
                  prior = priors,
                  sample_prior = 'only',
                  iter = 5000,
                  warmup = 1000,
                  chains = 3,
                  thin = 5,
                  refresh = 0
                  )
```

```{r partialPlot2h1a, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
growth.brm %>%
  conditional_effects() |>
  plot(points = TRUE)
```

```{r fitModel2h1b, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
growth.brm2 <- update(growth.brm,  
                       sample_prior = 'yes',
                       control = list(adapt_delta = 0.99),
                       refresh = 0)
#save(growth.brm2, file = 'growth.brm2')
```

```{r}
growth.brm2 %>%
    conditional_effects() %>%
    plot(points = TRUE)
```

### Plotting prior and posterior
```{r}
growth.brm2 %>% SUYR_prior_and_posterior()
```

```{r}
growth.brm2$fit |> stan_trace()
```

```{r}
growth.brm2$fit |> stan_ac()
```

```{r}
growth.brm2$fit |> stan_rhat()
```

```{r}
growth.brm2$fit |> stan_ess()
```

### Random intercept/slope model
```{r fitModel2h3, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
#priors<- prior("normal(-10, 5)", class = "Intercept") +
#    prior("normal(0, 15)", class = "b") +
#    #prior("normal(0, 0.5)", class = "sd", group = "Tank") +
#    prior("normal(0, 20)", class = "sd", group = "Plate.Slide") +
#    prior(student_t(3, 0, 24.6), class = 'sigma') +
#    prior(cauchy(0,2.5), class = 'sd') +
#    prior(lkj_corr_cholesky(1), class = 'L')
```

```{r}
#growth.form <- bf(Growth ~ popcross * Temp + (popcross|Plate.Slide),
#                     family = gaussian()
#                   )
```

```{r}
#growth.brm3 <- brm(growth.form, 
#                  data = growth,
#                  prior = priors,
#                  sample_prior = 'yes',
#                  iter = 5000,
#                  warmup = 1000,
#                  chains = 3,
#                  thin = 5,
#                  refresh = 0,
#                  control = list(adapt_delta = 0.99)
#                  )
#save(growth.brm3, file = 'growth.brm3')
```

```{r}
#growth.brm3 %>%
#    conditional_effects() %>%
#    plot(points = TRUE)
```

### Plotting prior and posterior
```{r}
#growth.brm3%>% SUYR_prior_and_posterior()
```

```{r}
#growth.brm3$fit |> stan_trace()
```

```{r}
#growth.brm3$fit |> stan_ac()
```

```{r}
#growth.brm3$fit |> stan_rhat()
```

```{r}
#growth.brm3$fit |> stan_ess()
```

### Compare models
```{r fitModel2h3a, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
#loo(growth.brm2, growth.brm3)
```

# Model validation
```{r modelValidation5b, results='markdown', eval=TRUE, hidden=TRUE, fig.width=6, fig.height=4}
growth.brm2 %>% pp_check(type = 'dens_overlay', nsamples = 250)
```

```{r modelValidation5c, results='markdown', eval=TRUE, hidden=TRUE, fig.width=6, fig.height=4}
growth.brm2 %>% pp_check(type = 'error_scatter_avg')
```

### DHARMa residuals
```{r}
growth.resids<- make_brms_dharma_res(growth.brm2, integerResponse = FALSE)
wrap_elements(~testUniformity(growth.resids)) +
wrap_elements(~plotResiduals(growth.resids, form = factor(rep(1,nrow(growth))))) +
wrap_elements(~plotResiduals(growth.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(growth.resids))
```

**Conclusions:**
### conditional_effects
```{r partialPlot2d, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
growth.brm2 %>%
    conditional_effects("popcross:Temp") %>%
    plot(points = TRUE)
```

### ggpredict
```{r partialPlot2a, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
growth.brm2 %>%
    ggpredict(~popcross*Temp) %>%
    plot(add.data = TRUE)
```

### ggemmeans
```{r partialPlot2b, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
starling.brm3 %>%
    ggemmeans(~popcross*Temp) %>%
    plot(add.data = TRUE) 
```

```{r}
growth.brm2 |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 0),
    Pg = ~mean(.x > 0),
    rhat,
    ess_bulk
    ) |>
  knitr::kable()
```

```{r}
growth.brm2 |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 0),
    Pg = ~mean(.x > 0),
    rhat,
    ess_bulk
    ) |>
  knitr::kable()
```

```{r summariseModel2c4, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
growth.brm2 %>%
    gather_draws(`b_Intercept.*|b_popcross.*|b_Temp.*`, regex=TRUE) %>% 
    ggplot() +
    geom_vline(xintercept=0, linetype='dashed') +
    stat_slab(aes(x = .value, y = .variable,
                  fill = stat(ggdist::cut_cdf_qi(cdf,
                                                 .width = c(0.5, 0.8, 0.95), 
                                                 labels = scales::percent_format())
                              )), color='black') + 
    scale_fill_brewer('Interval', direction = -1, na.translate = FALSE) 
```

```{r}
growth.brm2 %>% 
    gather_draws(`.Intercept.*|b_popcross.*|b_Temp.*`, regex=TRUE) %>% 
    ggplot() + 
    geom_vline(xintercept = 0, linetype='dashed') +
    stat_halfeye(aes(x=.value,  y=.variable)) +
    theme_classic()
```

### density ridges (ggridges)
```{r summariseModel2c7, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
##in colour
growth.brm2 %>% 
    gather_draws(`^b_.*`, regex=TRUE) %>% 
    filter(.variable != 'b_Intercept') %>%
    ggplot() + 
    geom_density_ridges_gradient(aes(x=(.value),
                                     y = .variable,
                                     fill = stat(x)),
                                 alpha=0.4, colour = 'white',
                                 quantile_lines = TRUE,
                                 quantiles = c(0.025, 0.975)) +
    geom_vline(xintercept = 1, linetype = 'dashed') +
    scale_x_continuous() +
    scale_fill_viridis_c(option = "C") 

## Fractional scale
growth.brm2 %>% 
    gather_draws(`^b_.*`, regex=TRUE) %>% 
    filter(.variable != 'b_Intercept') %>%
    ggplot() + 
    geom_density_ridges_gradient(aes(x=exp(.value),
                                     y = .variable,
                                     fill = stat(x)),
                                 alpha=0.4, colour = 'white',
                                 quantile_lines = TRUE,
                                 quantiles = c(0.025, 0.975)) +
    geom_vline(xintercept = 1, linetype = 'dashed') +
    scale_x_continuous(trans = scales::log2_trans()) +
    scale_fill_viridis_c(option = "C") 
```
 
### R2
```{r summariseModel2g, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
growth.brm2 %>%
    bayes_R2(re.form = NA, summary=FALSE) %>%
    median_hdci
growth.brm2 %>%
    bayes_R2(re.form = ~(1|Plate.Slide), summary=FALSE) %>%
    median_hdci
growth.brm2 %>%
    bayes_R2(re.form = ~(popcross|Plate.Slide), summary=FALSE) %>%
    median_hdci
```

```{r}
growth.brm2 |>
  emmeans(~popcross, type = "response")
```

planned contrasts- compare situations

              North vs Central    Central vs Hybrid   North vs hybrid  N mom vs N dad
              
DAVxDAV     |    -1                -1                     0                 0  
DAVxMAR     |     0                 1/4                   1/4               1/3     
MARxARL     |     0                 1/4                   1/4              -1/3     
MARxDAV     |     0                 1/4                   1/4               1/3    
MARxMAR     |     1/3               0                    -1/3               0    
MARxWOO     |     1/3               0                    -1/3               0    
PALxMAR     |     0                 1/4                   1/4               1/3     
WOOxMAR     |     1/3               0                    -1/3               0     

```{r}
cmat<- cbind("North vs Central"=c(-1, 0, 0, 0, 1/3, 1/3, 0, 1/3),
             "Central vs Hybrid" = c(-1, 1/4, 1/4, 1/4, 0, 0, 1/4, 0),
             "North vs hybrid" = c(0, 1/4, 1/4, 1/4, -1/3, -1/3, 1/4, -1/3),
             "N mom vs N dad" = c(0, 1/3, -1/3, 1/3, -0, 0, 1/3, 0)
             )
```
 
```{r}
growth.brm2 |>
  emmeans(~popcross) |>
  contrast(list(popcross=cmat)) |>
    gather_emmeans_draws() |>
  summarise(median_hdci(.value),
            Pl = mean(.value < 0),
            Pg = mean(.value > 0)
            )
```

```{r}
growth.brm2 |>
  emmeans(~popcross|Temp) |>
  as.data.frame() |>
  ggplot(aes(y = emmean, x = popcross, color = Temp)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD))
```

#############################################################################################################
Hot only
```{r}
growth.hot<- growth |>
  filter(Temp == "HOT")
```

```{r}
growth.form <- bf(Growth ~ popcross + (1|Plate.Slide),
                     family = gaussian()
                   )
```

```{r fitModel2h1, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
priors<- prior("normal(-10, 5)", class = "Intercept") +
    prior("normal(0, 15)", class = "b") +
    prior("normal(0, 7)", class = "sd", group = "Plate.Slide") +
    prior(student_t(2, -5, 1), class = 'sigma') +
    prior(cauchy(0,2.5), class = 'sd')
```

```{r}
growth.brm4 <- brm(growth.form, 
                  data = growth.hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 1000,
                  chains = 3,
                  thin = 5,
                  control = list(adapt_delta = 0.99),
                  refresh = 0
                  )
```

```{r partialPlot2h1a, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
growth.brm4 %>%
  conditional_effects() |>
  plot(points = TRUE)
```

### Plotting prior and posterior
```{r}
growth.brm4 %>% SUYR_prior_and_posterior()
```

```{r}
growth.brm4$fit |> stan_trace()
```

```{r}
growth.brm4$fit |> stan_ac()
```

```{r}
growth.brm4$fit |> stan_rhat()
```

```{r}
growth.brm4$fit |> stan_ess()
```

# Model validation
```{r modelValidation5b, results='markdown', eval=TRUE, hidden=TRUE, fig.width=6, fig.height=4}
growth.brm4 %>% pp_check(type = 'dens_overlay', nsamples = 250)
```

### DHARMa residuals
```{r}
growth.resids<- make_brms_dharma_res(growth.brm4, integerResponse = FALSE)
wrap_elements(~testUniformity(growth.resids)) +
wrap_elements(~plotResiduals(growth.resids, form = factor(rep(1,nrow(growth.hot))))) +
wrap_elements(~plotResiduals(growth.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(growth.resids))
```

```{r summariseModel2c4, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
growth.brm4 %>%
    gather_draws(`b_Intercept.*|b_popcrossxsym.*`, regex=TRUE) %>% 
    ggplot() +
    geom_vline(xintercept=0, linetype='dashed') +
    stat_slab(aes(x = .value, y = .variable,
                  fill = stat(ggdist::cut_cdf_qi(cdf,
                                                 .width = c(0.5, 0.8, 0.95), 
                                                 labels = scales::percent_format())
                              )), color='black') + 
    scale_fill_brewer('Interval', direction = -1, na.translate = FALSE) 
```

####################################################################################################
popcross only
```{r}
growth.form <- bf(Growth ~ popcross + (1|Plate.Slide),
                     family = gaussian()
                   )
```

### Defining priors
```{r fitModel2h1, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
priors<- prior("normal(-10, 5)", class = "Intercept") +
    prior("normal(0, 15)", class = "b") +
    #prior("normal(0, 0.5)", class = "sd", group = "Tank") +
    prior("normal(0, 7)", class = "sd", group = "Plate.Slide") +
    prior(student_t(2, -5, 1), class = 'sigma') +
    prior(cauchy(0,2.5), class = 'sd')
```

```{r}
growth.brm5 <- brm(growth.form, 
                  data = growth.hot,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 5000,
                  warmup = 1000,
                  chains = 3,
                  thin = 5,
                  refresh = 0
                  )
```

```{r partialPlot2h1a, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
growth.brm5 %>%
  conditional_effects() |>
  plot(points = TRUE)
```

```{r summariseModel2c4, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
growth.brm5 %>%
    gather_draws(`b_Intercept.*|b_popcross.*`, regex=TRUE) %>% 
    ggplot() +
    geom_vline(xintercept=0, linetype='dashed') +
    stat_slab(aes(x = .value, y = .variable,
                  fill = stat(ggdist::cut_cdf_qi(cdf,
                                                 .width = c(0.5, 0.8, 0.95), 
                                                 labels = scales::percent_format())
                              )), color='black') + 
    scale_fill_brewer('Interval', direction = -1, na.translate = FALSE) 
```