---
title: "Physiology- A. tenuis juvenile YII"
author: "Alex Macadam"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the necessary libraries
```{r libraries, results='markdown', eval=TRUE, message=FALSE, warning=FALSE}
library(car)       #for regression diagnostics
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(sjPlot)    #for outputs
library(knitr)     #for kable
library(effects)   #for partial effects plots
library(emmeans)   #for estimating marginal means
library(MASS)      #for glm.nb
library(MuMIn)     #for AICc
library(tidyverse) #for data wrangling
library(brms)
library(tidybayes)
library(broom.mixed)#for tidying MCMC outputs
library(patchwork)  #for multiple plots
library(rstanarm)
library(ggeffects)
library(bayesplot)
library(rstan)
library(DHARMa)
library(ggridges)
source('/Users/alexmacadam/Dropbox/PhD/Chapter5_Outplanting/Chapter 5 Scripts/Chapter-5--Outplanting-AGF-corals/helperFunctions.R')
```

```{r}
# Function for summarising survival standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

# Read in the data
```{r readData, results='markdown', eval=TRUE}
yii <- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter2_Physiology/Chapter 2 data/Bayes stats data/physiology/juv_PAM2020.csv", header=T, fileEncoding="UTF-8-BOM")
```

# Exploratory data analysis
```{r processData, results='markdown', eval=TRUE}
#change catagorical columns to factors and remove NAs
yii <- yii |>
  mutate(Family = factor(Family),
         Species = factor(Species),
         Temp = factor(Temp),
         Tank = factor(Tank),
         Timepoint = factor(Timepoint),
         Cassette = factor(Cassette),
         Plate.Slide = factor(Plate.Slide),
         Treatment = factor(Treatment),
         Cone = factor(Cone),
         Crosstype = factor(Crosstype),
         popcross = factor(popcross),
         SINGLE.CLUMP = factor(SINGLE.CLUMP),
         Mom = factor(Mom),
         Dad = factor(Dad)
         ) |>
  filter(Crosstype == "IndivCross", SINGLE.CLUMP == "S", Timepoint %in% c("T-2","T6"))
```

Add Temp x Treatment column
```{r}
yii$symxtemp <- paste(yii$Treatment, yii$Temp, sep='_')
yii$symxtemp <- as.factor(yii$symxtemp)
```

```{r}
yii2<- yii |>
  dplyr::select(Treatment, Timepoint, Temp,  YII_micro, IndID, popcross, Tank, Cassette, Plate.Slide, symxtemp) |>
  mutate(popcross = factor(popcross)) |>
  pivot_wider(names_from = Timepoint, values_from = YII_micro) |>
  tidyr::unnest()
```

```{r}
yii2<- yii2 |>
  mutate(YII_change = (`T6` - `T-2`)*1000) |> #times 1000 to make bigger numbers
  drop_na(YII_change)
```

```{r}
yii2$`T-2`[yii2$`T-2`<0.001] <- 0.001
yii2$T6[yii2$T6<0.001] <- 0.001
```

```{r}
yii2<- yii2 |>
  mutate(YII_p_change = ((`T6` - `T-2`)/`T-2`)*100) |>
  drop_na(YII_p_change)

#yii2<- yii2 |>
#  mutate(T6 = T6 *1000) |>
#  mutate(`T-2` = `T-2` * 1000) |>
#  mutate(YII_change = `T6` - `T-2`) |>
#  drop_na(YII_change)
```

Visualise
```{r}
ggplot(yii2, aes(y = YII_change, x = popcross, colour = Temp)) +
  geom_point(position = position_jitter(width = 0.2, height = 0)) +
  facet_wrap(~Treatment)

ggplot(yii2, aes(y = YII_change, x = popcross, colour = Temp)) +
  geom_boxplot() +
  facet_wrap(~Treatment)


```

```{r}
yii<- yii |>
  filter(Timepoint == "T6") |>
  mutate(YII_micro = YII_micro * 100)

yii |>
  ggplot(aes(y = YII_micro, x = popcross, colour = Temp)) +
  geom_point(position = position_jitter(width = 0.2, height = 0)) +
  facet_wrap(~Treatment)
```

Sumarise final YII
```{r}
yii.fin.sum.all.se <- summarySE(yii, measurevar="YII_micro",
                          groupvars=c("Temp"), na.rm=TRUE)

yii.fin.sum.se <- summarySE(yii, measurevar="YII_micro",
                          groupvars=c("Temp", "Treatment"), na.rm=TRUE)
```

Sumarise change in YII
```{r}
yii.sum.se <- summarySE(yii2, measurevar="YII_change",
                          groupvars=c("Temp", "Treatment"), na.rm=TRUE)
```

#####################################################################
Final YII

### Using default priors
```{r}
yii.form <- bf(YII_micro ~ symxtemp + (1|Plate.Slide),
                     family = gaussian()
                   )
```

```{r}
options(width=150)
yii.form %>% get_prior(data = yii)
options(width=80)
```

### Defining priors
```{r}
yii |>
group_by(Temp, Treatment) |>
summarise(
mean(YII_micro),
median(YII_micro),
sd(YII_micro),
mad(YII_micro)
)
```

```{r fitModel2h1, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
priors<- prior("normal(10, 20)", class = "Intercept") +
    prior("normal(0, 20)", class = "b") +
    #prior("normal(0, 0.5)", class = "sd", group = "Tank") +
    prior("normal(0, 20)", class = "sd", group = "Plate.Slide") +
    prior(student_t(3, 0, 20), class = 'sigma') +
    prior(cauchy(0,2.5), class = 'sd')

```

```{r}
YII.brm2 <- brm(yii.form, 
                  data = yii,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 10000,
                  warmup = 5000,
                  chains = 4,
                  thin = 20,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0
                  )
```

```{r}
#saveRDS(YII.brm2, file = "YII.brm2_A.tenuis_juv.rds")
YII.brm2 <- readRDS("YII.brm2_A.tenuis_juv.rds")
```

### Plotting prior and posterior
```{r}
YII.brm2 %>% SUYR_prior_and_posterior()
```

# Model validation
```{r}
YII.brm2$fit |> stan_trace()
```

```{r}
YII.brm2$fit |> stan_ac()
```

```{r}
YII.brm2$fit |> stan_rhat()
```

```{r}
YII.brm2$fit |> stan_ess()
```

```{r modelValidation5b, results='markdown', eval=TRUE, hidden=TRUE, fig.width=6, fig.height=4}
YII.brm2 %>% pp_check(type = 'dens_overlay', nsamples = 250)
```

### DHARMa residuals
```{r}
YII.resids<- make_brms_dharma_res(YII.brm2, integerResponse = FALSE)
wrap_elements(~testUniformity(YII.resids)) +
wrap_elements(~plotResiduals(YII.resids, form = factor(rep(1,nrow(YII))))) +
wrap_elements(~plotResiduals(YII.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(YII.resids))
```

### Compare models
```{r fitModel2h3a, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
#loo(YII.brm2, YII.brm3)
```

**Conclusions:**
### conditional_effects
```{r partialPlot2d, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
YII.brm2 %>%
    conditional_effects("symxtemp") %>%
    plot(points = TRUE)
```

##Summarise draws
```{r}
YII.brm2 |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 0),
    Pg = ~mean(.x > 0),
    rhat,
    ess_bulk
    ) |>
  knitr::kable()
```

```{r summariseModel2c4, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
YII.brm2 %>%
    gather_draws(`b_Intercept.*|b_symxtemp.*`, regex=TRUE) %>% 
    ggplot() +
    geom_vline(xintercept=0, linetype='dashed') +
    stat_slab(aes(x = .value, y = .variable,
                  fill = stat(ggdist::cut_cdf_qi(cdf,
                                                 .width = c(0.5, 0.8, 0.95), 
                                                 labels = scales::percent_format())
                              )), color='black') + 
    scale_fill_brewer('Interval', direction = -1, na.translate = FALSE) 
```

### density ridges (ggridges)
```{r summariseModel2c7, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
##in colour
YII.brm2 %>% 
    gather_draws(`^b_.*`, regex=TRUE) %>% 
    filter(.variable != 'b_Intercept') %>%
    ggplot() + 
    geom_density_ridges_gradient(aes(x=(.value),
                                     y = .variable,
                                     fill = stat(x)),
                                 alpha=0.4, colour = 'white',
                                 quantile_lines = TRUE,
                                 quantiles = c(0.025, 0.975)) +
    geom_vline(xintercept = 1, linetype = 'dashed') +
    scale_x_continuous() +
    scale_fill_viridis_c(option = "C") 
```
 
### R2
```{r summariseModel2g, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
YII.brm2 %>%
    bayes_R2(re.form = ~(1|Plate.Slide), summary=FALSE) %>%
    median_hdci
```


#Emmeans
```{r}
YII.brm2 |>
  emmeans(~symxtemp, type = "response")
```
planned contrasts- compare situations

                 D1 vs Sed       D1 vs SS        SS vs Sed       Central vs Hybrid       
D1_AMB     |      0                -1/2                0                   -1/3
D1_HOT     |     -1                -1/2                0                    1/4
Jewell_AMB |      0              0                  1/2                 -1/3
Jewell_HOT |      0              0                  1/2                  1/4
SS_AMB     |      0                1/2               -1/2                 -1/3
SS_HOT     |      1                1/2               -1/2                  1/4
 
```{r}
cmat<- cbind(
             "D1_HOT vs Sed_HOT"=c(0, -1, 0, 1, 0, 0),
             "D1_HOT vs SS_HOT" = c(0, -1, 0, 0, 0, 1),
             "SS_HOT vs Sed_HOT" = c(0, 0, 0, 1, 0, -1),
             "D1_AMB vs Sed_AMB"=c(-1, 0, 1, 0, 0, 0),
             "D1_AMB vs SS_AMB" = c(-1, 0, 0, 0, 1, 0),
             "SS_AMB vs Sed_AMB" = c(0, 0, 1, 0, -1, 0),
             "D1_AMB vs D1_HOT" = c(1, -1, 0, 0, 0, 0),
             "SS_AMB vs SS_HOT" = c(0, 0, 0, 0, 1, -1),
             "SED_AMB vs SED_HOT" = c(0, 0, 1, -1, 0, 0)
             )
```
 
```{r}
YII.brm2 |>
  emmeans(~symxtemp) |>
  contrast(list(symxtemp=cmat)) |>
  gather_emmeans_draws() |>
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 0),
            Pg = mean(.value > 0)
            )
```

```{r}
YII.brm2 |>
  emmeans(~symxtemp) |>
  as.data.frame() |>
  ggplot(aes(y = emmean, x = symxtemp)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD))
```

```{r}
YII.plot<- YII.brm2 %>%
  emmeans(~symxtemp) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = symxtemp,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Final YII", y = "Treatment")
YII.plot
#saveRDS(YII.plot, file = "YII.plot_A.tenuis_juv.rds")
```

#############################################################################################################
YII change

### Using default priors
```{r}
yii.form <- bf(YII_change ~ symxtemp + (1|Plate.Slide),
                     family = gaussian()
                   )
```

```{r}
options(width=150)
yii.form %>% get_prior(data = yii2)
options(width=80)
```

### Defining priors
```{r}
yii2 |>
group_by(Temp, Treatment) |>
summarise(
mean(YII_change),
median(YII_change),
sd(YII_change),
mad(YII_change)
)
```

```{r fitModel2h1, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
priors<- prior("normal(1000, 5000)", class = "Intercept") +
    prior("normal(1000, 5000)", class = "b") +
    #prior("normal(0, 0.5)", class = "sd", group = "Tank") +
    prior("normal(0, 5000)", class = "sd", group = "Plate.Slide") +
    prior(student_t(3, 0, 2000), class = 'sigma') +
    prior(cauchy(0,2.5), class = 'sd')

priors<- prior("normal(0, 200)", class = "Intercept") +
    prior("normal(0, 200)", class = "b") +
    #prior("normal(0, 0.5)", class = "sd", group = "Tank") +
    prior("normal(0, 300)", class = "sd", group = "Plate.Slide") +
    prior(student_t(3, 0, 100), class = 'sigma') +
    prior(cauchy(0,2.5), class = 'sd')
```

```{r}
YII.brm3 <- brm(yii.form, 
                  data = yii2,
                  prior = priors,
                  sample_prior = 'yes',
                  iter = 10000,
                  warmup = 5000,
                  chains = 4,
                  thin = 20,
                  control = list(adapt_delta = 0.99, max_treedepth = 20),
                  refresh = 0
                  )
```

```{r}
#saveRDS(YII.brm3, file = "YII.brm3_A.tenuis_juv.rds")
YII.brm3 <- readRDS("YII.brm3_A.tenuis_juv.rds")
```

### Plotting prior and posterior
```{r}
YII.brm3 %>% SUYR_prior_and_posterior()
```

# Model validation
```{r}
YII.brm3$fit |> stan_trace()
```

```{r}
YII.brm3$fit |> stan_ac()
```

```{r}
YII.brm3$fit |> stan_rhat()
```

```{r}
YII.brm3$fit |> stan_ess()
```

```{r modelValidation5b, results='markdown', eval=TRUE, hidden=TRUE, fig.width=6, fig.height=4}
YII.brm3 %>% pp_check(type = 'dens_overlay', nsamples = 250)
```

### DHARMa residuals
```{r}
YII.resids<- make_brms_dharma_res(YII.brm3, integerResponse = FALSE)
wrap_elements(~testUniformity(YII.resids)) +
wrap_elements(~plotResiduals(YII.resids, form = factor(rep(1,nrow(YII))))) +
wrap_elements(~plotResiduals(YII.resids, quantreg = FALSE))+
wrap_elements(~testDispersion(YII.resids))
```

### Compare models
```{r fitModel2h3a, results='markdown', eval=TRUE, hidden=TRUE, cache=TRUE}
#loo(YII.brm3, YII.brm3)
```

**Conclusions:**
### conditional_effects
```{r partialPlot2d, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
YII.brm3 %>%
    conditional_effects("symxtemp") %>%
    plot(points = TRUE)
```

##Summarise draws
```{r}
YII.brm3 |>
  summarise_draws(
    median,
    HDInterval::hdi,
    Pl = ~mean(.x < 0),
    Pg = ~mean(.x > 0),
    rhat,
    ess_bulk
    ) |>
  knitr::kable()
```

```{r summariseModel2c4, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
YII.brm3 %>%
    gather_draws(`b_Intercept.*|b_symxtemp.*`, regex=TRUE) %>% 
    ggplot() +
    geom_vline(xintercept=0, linetype='dashed') +
    stat_slab(aes(x = .value, y = .variable,
                  fill = stat(ggdist::cut_cdf_qi(cdf,
                                                 .width = c(0.5, 0.8, 0.95), 
                                                 labels = scales::percent_format())
                              )), color='black') + 
    scale_fill_brewer('Interval', direction = -1, na.translate = FALSE) 
```

### density ridges (ggridges)
```{r summariseModel2c7, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5,echo=TRUE}
##in colour
YII.brm3 %>% 
    gather_draws(`^b_.*`, regex=TRUE) %>% 
    filter(.variable != 'b_Intercept') %>%
    ggplot() + 
    geom_density_ridges_gradient(aes(x=(.value),
                                     y = .variable,
                                     fill = stat(x)),
                                 alpha=0.4, colour = 'white',
                                 quantile_lines = TRUE,
                                 quantiles = c(0.025, 0.975)) +
    geom_vline(xintercept = 1, linetype = 'dashed') +
    scale_x_continuous() +
    scale_fill_viridis_c(option = "C") 
```
 
### R2
```{r summariseModel2g, results='markdown', eval=TRUE, hidden=TRUE, fig.width=8, fig.height=5}
YII.brm3 %>%
    bayes_R2(re.form = ~(1|Plate.Slide), summary=FALSE) %>%
    median_hdci
```


#Emmeans
```{r}
YII.brm3 |>
  emmeans(~symxtemp, type = "response")
```
planned contrasts- compare situations

                 D1 vs Sed       D1 vs SS        SS vs Sed       Central vs Hybrid       
D1_AMB     |      0                -1/2                0                   -1/3
D1_HOT     |     -1                -1/2                0                    1/4
Jewell_AMB |      0              0                  1/2                 -1/3
Jewell_HOT |      0              0                  1/2                  1/4
SS_AMB     |      0                1/2               -1/2                 -1/3
SS_HOT     |      1                1/2               -1/2                  1/4
 
```{r}
cmat<- cbind(
             "D1_HOT vs Sed_HOT"=c(0, -1, 0, 1, 0, 0),
             "D1_HOT vs SS_HOT" = c(0, -1, 0, 0, 0, 1),
             "SS_HOT vs Sed_HOT" = c(0, 0, 0, 1, 0, -1),
             "D1_AMB vs Sed_AMB"=c(-1, 0, 1, 0, 0, 0),
             "D1_AMB vs SS_AMB" = c(-1, 0, 0, 0, 1, 0),
             "SS_AMB vs Sed_AMB" = c(0, 0, 1, 0, -1, 0),
             "D1_AMB vs D1_HOT" = c(1, -1, 0, 0, 0, 0),
             "SS_AMB vs SS_HOT" = c(0, 0, 0, 0, 1, -1),
             "SED_AMB vs SED_HOT" = c(0, 0, 1, -1, 0, 0)
             )
```
 
```{r}
YII.brm3 |>
  emmeans(~symxtemp) |>
  contrast(list(symxtemp=cmat)) |>
  gather_emmeans_draws() |>
  dplyr::summarise(median_hdci(.value),
            Pl = mean(.value < 0),
            Pg = mean(.value > 0)
            )
```

```{r}
YII.brm3 |>
  emmeans(~symxtemp) |>
  as.data.frame() |>
  ggplot(aes(y = emmean, x = symxtemp)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD))
```

```{r}
YII.change.plot<- YII.brm3 %>%
  emmeans(~symxtemp) %>%
  gather_emmeans_draws() %>%
    ggplot() +
  #geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(aes(
    x = .value, y = symxtemp,
    fill = stat(ggdist::cut_cdf_qi(cdf,
      .width = c(0.5, 0.8, 0.95),
      labels = scales::percent_format()
    ))
  ), color = "black") +
  scale_fill_brewer("Interval", direction = -1, na.translate = FALSE) +
  theme_classic() +
  labs(x = "Change in YII", y = "Treatment")
YII.change.plot
#saveRDS(YII.change.plot, file = "YII.change.plot_A.tenuis_juv.rds")
```
